Classification of ordinal data in deep learning: an experimental study

### 2019-02-27 paper/ieee.tex
paper

@@

% *** PDF, URL AND HYPERLINK PACKAGES ***
%
\usepackage{url}
\usepackage{hyperref}
% url.sty was written by Donald Arseneau. It provides better support for
% handling and breaking URLs. url.sty is already installed on most LaTeX
% systems. The latest version and documentation can be obtained at:
% http://www.ctan.org/pkg/url
% Basically, \url{my_url_here}.

@@

@@

% correct bad hyphenation here
\hyphenation{op-tical net-works semi-conduc-tor}

@@

	\author{Victor-Manuel~Vargas, Pedro-Antonio~Gutiérrez and César~Hervás}

@@

	\section{Introduction}
	\label{sect:introduction}
	% The very first letter is a 2 line initial drop letter followed
	% by the rest of the first word in caps.
	% 
	% form to use if the first word consists of a single letter:
	% \IEEEPARstart{A}{demo} file is ....
	% 
	% form to use if you need the single drop letter followed by
	% normal text (unknown if ever used by the IEEE):
	% \IEEEPARstart{A}{}demo file is ....
	% 
	% Some journals put the first two words in caps:
	% \IEEEPARstart{T}{his demo} file is ....
	% 
	% Here we have the typical use of a "T" for an initial drop letter
	% and "HIS" in caps to complete the first word.
	\IEEEPARstart{D}{eep learning}, introduced by Yann Lecun~\cite{lecun2015deep}, combines multiple Machine Learning techniques and allows computational models that are composed of numerous processing layers to learn representations of data with various levels of abstraction. These methods have dramatically improved the state-of-the-art in many domains, such as image classification~\cite{cirecsan2012multi,he2016deep,krizhevsky2012imagenet}, speech recognition~\cite{hinton2012deep}, control problems~\cite{mnih2015human}, object detection~\cite{jiang2016speed,girshick2014rich}, privacy protection~\cite{yu2017iprivacy}, recovery of human pose~\cite{hong2015multimodal}, semantic segmentation~\cite{long2015fully} and image retrieval~\cite{li2015weakly}. Deep learning discovers complex structures in large datasets by using the backpropagation algorithm to indicate how a machine should change its internal parameters that are used to compute the representation in each layer from the features in the previous layer.

	Convolutional Neural Networks (CNN) are designed to process data that comes in the form of multiple arrays. CNNs are suited for images, video, speech and audio processing, and they have been used extensively in the last years for automatic classification tasks~\cite{dong2014learning,sun2013deep,ronneberger2015u}. On image classification tasks, each colour channel is represented by a 2D array. In this case, convolutional layers extract the main features from the pixels of the images and, after that, a fully connected layer classify every sample based on its extracted features. At the output of the convolutional net, a softmax function provides the probabilities of the set of classes predefined in the model for classification tasks. These outputs are compared against the correct values.

	The backpropagation algorithm is a stochastic gradient descent algorithm that minimises a predefined loss function. It has been used in the last years in many works~\cite{leonard1990improvement,yu1995dynamic,krizhevsky2012imagenet,de2018weighted} for training shallow and deep neural networks. It updates the layer's parameters after backpropagating the loss function gradients through the network. Learning rate hyper-parameter controls the strength of the changes that are applied to those parameters. Some works have checked the importance of finding the optimal value for this parameter~\cite{senior2013empirical} and have tried different approaches to try to improve the training process~\cite{smith2017cyclical}.

	Batch normalization is another technique that is used for this kind of networks. It reduces the internal covariate shift by normalizing layer inputs. It was presented in 2015~\cite{ioffe2015batch} and gives a critical enhancement to the training phase. Batch size is an important hyper-parameter that should be adjusted as it affects the layer's parameters updates and also the normalization~\cite{keskar2016large}\cite{radiuk2017impact}.

	Ordinal classification problems are those where labels are ordered, and there are different inter-classes weights for each pair of classes. This kind of problem can be treated as a nominal classification problem, but when you do this, you are discarding ordinal information. A better approach is to use specific methods that take into account this kind of information to improve the performance of the classification model. One way to use the ordinal information is to evaluate the model using an ordinal metric. Multiple metrics exist in the literature of machine learning and statistics \cite{cruz2014metrics,mehdiyev2016evaluating}. Kappa index is a well-known statistic coefficient defined by Cohen \cite{cohen1960coefficient} to measure inter-rater agreement on classifying elements into a set of categories. Later, Weighted Kappa (WK) is a modified version of the Kappa statistic calculated to allow assigning different weights to different levels of aggregation between two variables. Weighted kappa loss was described in previous work~\cite{de2018weighted} and is a cost function that is based on the WK metric. These functions are indicated for problems where different inter-classes weights are assigned, and that work proved that using these functions improves the model performance and reduce the overfitting risk. This kind of problems is associated with ordinal data processing. Those weights are predefined and depend on the type of the chosen WK. In a linear penalization, the weight is proportional to the distance between the predicted and the real class. In the Quadratic Weighted Kappa (QWK) the penalization is proportional to the square of the distance. In this work, we will use the QWK metric and the QWK loss.

	Softmax function is widely used for the output layer in neural networks for classification tasks. It is a simple and efficient function for multi-class problems but not the best when working with ordinal data. In this paper, the Proportional Odds Model (POM)~\cite{gutierrez2016ordinal,agresti2010analysis} will be used instead of the softmax function. Different link functions will be explored to compare their performance. Also, the influence of other parameters like the learning rate of the optimization algorithm and the batch size will be studied. We will contrast the results obtained with statistical analysis to provide more robust results. An approximated ANOVA III \cite{miller1997beyond} test followed by a posthoc Tukey's test \cite{tukey1949comparing} will be performed because of the limitations of the computational time required to run a higher number of executions.

	The experiments will be run using two different ordinal datasets: Diabetic Retinopathy~\cite{de2018weighted}, that contains high-resolution fundus images related with diabetes disease, and Adience~\cite{beckham2017unimodal}, that is formed of human faces images that are assigned an age range.

	This paper is organized as follows: in Section~\ref{sect:relatedwork}, we take a look at previous works related to this paper. Sections~\ref{sect:ordinalproblem} and \ref{sect:pom} present a formal description of an ordinal problem and the Proportional Odds Model. In Section~\ref{sect:experiments} we describe the model, the experiments and the datasets used, in Section~\ref{sect:results} we present the results obtained and the statistical analysis and finally in Section~\ref{sect:conclusions} we expose the conclusions of this work.

	\section{Related work}
	\label{sect:relatedwork}
	There are many works related to the application and development of convolutional neural networks models. Few works treat ordinal classification problems, though.

	J. de la Torre et al.~\cite{de2018weighted} proposed the use of QWK loss function for the optimization algorithm. They compared this cost function against the traditional log-loss function across three different databases, including the Diabetic Retinopathy database as the most complex one. They proved that their function could improve the results as it reduces the overfitting and the required training time. Also, they checked the importance of hyper-parameter tuning.

	Christopher Beckham and Christopher Pal~\cite{beckham2017unimodal} proposed a straightforward technique to constrain discrete ordinal probability distributions to be unimodal via the use of the Poisson and binomial probability distributions. They evaluated this approach in the context of deep learning on two large ordinal image datasets, including the Adience dataset used in this paper, and they obtained promising results.

	Adience dataset has been used in other works for human age estimation. E. Eidinger \cite{eidinger2014age} presented an approach using support vector machines and neural networks. J.-C. Chen \cite{chen2016cascaded} proposed a coarse-to-fine strategy for deep convolutional networks. G. Levi \cite{levi2015age} presented another convolutional network model for age estimation.

	At present, there is a great discussion in terms of finding the best activation function that offers good performance and mitigates the effects of the gradients vanishing problem \cite{bengio1994learning,pascanu2013difficulty}. Rectified Linear Unit (ReLU)~\cite{nair2010rectified} is widely used in most deep learning works, but recently, Clevert et al. proposed the Exponential Linear Unit (ELU)~\cite{clevert2015fast}. They proved that ELUs alleviate the vanishing gradient problem via the identity for positive values. In their experiments, ELUs lead not only to faster learning but also to significantly better generalization performance than ReLUs and Leaky ReLU functions (LReLU) on networks with more than five layers. That's why we decided to use the ELU function for the experiments described in this paper.

	Sergey Ioffe and Christian Szegedy~\cite{ioffe2015batch} described Batch Normalization and its benefits in previous work. It reduces the internal covariate shift by normalizing layer inputs. Their method draws its strength from making normalization a part of the model architecture and performing the normalization for each training batch. It allows them to use higher learning rates and be less careful about initialization. It also eliminates the need for using regularization techniques like Dropout.

	\section{Ordinal problem definition}
	\label{sect:ordinalproblem}
	The ordinal problem consists of predicting the label y of an input vector $\mathbf{x}$, where $\mathbf{x} \in \mathcal{X} \subseteq \mathds{R}^K$ and $y \in \mathcal{Y}~=~\{\mathcal{C}_1, \mathcal{C}_2, ..., \mathcal{C}_Q\}$, i.e. $\mathbf{x}$ is in a K-dimensional input space, and $y$ is in a label space of $Q$ different labels. The objective of the ordinal problem is to find a function $r : \mathcal{X} \rightarrow \mathcal{Y}$ to predict the labels or categories of new patterns, given a training set of $N$ points, $D = \{(x_i, y_i), i = 1, ..., N\}$. A natural label ordering exists for ordinal problems: $\mathcal{C}_1 \prec \mathcal{C}_2 \prec ... \prec \mathcal{C}_Q$. The assumption of order between class labels makes that two different elements of $\mathcal{Y}$ can always be compared by using the relation $\prec$, which is not possible under the nominal classification setting. In regression (where $y \in \mathds{R}$), real values in $\mathds{R}$ can be ordered by the standard $<$ operator, but labels in ordinal regression ($y \in \mathcal{Y}$) do not carry metric information, so the category serves as a qualitative indication of the pattern rather than a quantitative one.

	\section{Proportional Odds Model}
	\label{sect:pom}
	Arising from a statistical background, the Proportional Odds Model (POM) is one of the first models designed explicitly for ordinal regression~\cite{mccullagh1980regression}, dated back to 1980. It is a member of a wider family of models recognised as Cumulative Link Models (CLMs)~\cite{agresti2010analysis}. CLMs predict probabilities of groups of contiguous categories, taking the ordinal scale into account. In this way, cumulative probabilities $P(y \prec C_q |\mathbf{x})$ are estimated, which can be directly related to standard probabilities:
	\begin{equation}
		P(y \preceq \mathcal{C}_q | \mathbf{x}) = P(y = \mathcal{C}_1 | \mathbf{x}) + ... + P(y = \mathcal{C}_q | \mathbf{x}),
	\end{equation}
	\begin{equation}
		P(y = \mathcal{C}_q | \mathbf{x}) = P(y \preceq \mathcal{C}_q | \mathbf{x}) - P(y \preceq \mathcal{C}_{q-1} | \mathbf{x}),
	\end{equation}
	with $q = 2, ..., Q-1$, and considering that $P(y = \mathcal{C}_1 | \mathbf{x}) = P(y \preceq \mathcal{C}_1 | \mathbf{x})$ and $P(y \preceq \mathcal{C}_Q | \mathbf{x}) = 1$.

	The model is inspired by the notion of latent variable, considering a linear transformation $f(x) = \mathbf{w^Tx}$ of the input variables as the one-dimensional mapping. The decision rule $r: \mathcal{X} \rightarrow \mathcal{Y}$ is not fitted directly, but stochastic ordering of space $\mathcal{X}$ is satisfied by the following general model form \cite{herbrich2000large}:
	\begin{equation}
		g^{-1}(P(y \preceq \mathcal{C}_q | \mathbf{x})) = b_q - \mathbf{w^Tx}, \quad q = 1, ..., Q-1,
	\end{equation}
	where $g^{-1} : [0,1] \rightarrow (-\infty, +\infty)$ is a monotonic function often termed as the inverse link function and $b_q$ is the threshold defined for class $\mathcal{C}_q$. Consider the latent variable $y^* = f(x)^* = \mathbf{w^Tx} + \epsilon$, where $\epsilon$ is the random variable of the error. The most common choice for the probability distribution of $\epsilon$ is the logistic function (which is the default function for POM). Label $\mathcal{C}_q$ is proposed if and only if $f(x) \in [b_{q-1}, b_q]$, where the function $f$ and $\mathbf{b} = (b_0, b_1, ..., b_{Q-1}, b_Q)$ are to be determined from the data. it is assumed that $b_0 = -\infty$ and $b_Q = +\infty$, so the real line defined by $f(x), x \in \mathcal{X}$, is divided into $Q$ consecutive intervals. Each interval corresponds to a category. The constraints $b_1 \le b_2 \le ... \le b_{Q-1}$ ensures that $P(y \preceq \mathcal{C}_q | \mathbf{x})$ increases with $q$ \cite{mccullagh1980regression}.
	In this work, we use this ordinal model with different link functions for the probability distribution of $\epsilon$, including logit, probit and complementary log-log. These functions will be detailed in Section \ref{sect:model}.
	\section{Experiments}
	\label{sect:experiments}
	\subsection{Data}
	We make use of two ordinal datasets appropriate for deep neural networks:

	\begin{itemize}
		\item \textit{Diabetic Retinopathy (DR)}\footnote{https://www.kaggle.com/c/diabetic-retinopathy-detection/data}. DR is a dataset consisting of extremely high-resolution fundus image data. The training set consists of $17563$ pairs of images (where a  pair consists of a left and right eye image corresponding to a patient). In this dataset, we try to predict from five levels of diabetic retinopathy: no DR ($25810$ images), mild DR ($2443$ images), moderate DR ($5292$ images), severe DR ($873$ images), or proliferative DR ($708$ images). The images are taken in variable conditions: by different cameras,  varying conditions of illumination and different resolutions. These images come from the EyePACS dataset that was used in a Diabetic Retinopathy Detection competition that was hosted on the Kaggle platform. Also, this dataset was used in later works~\cite{de2018weighted,nebot2016diabetic} and ordinal techniques (such as an ordinal cost function) were applied in order to achieve better performance. A validation set is set aside, consisting of $10\%$ of the patients in the training set. The images are resized to 128 by 128 pixels and rescaled to $[0,1]$ range. Data augmentation techniques, described in Section \ref{sect:procedure}, are applied to achieve a higher number of samples.

		\item \textit{Adience}\footnote{http://www.openu.ac.il/home/hassner/Adience/data.html}. This dataset consists of $26580$ faces belonging to $2284$ subjects. We use the form of the dataset where faces have been pre-cropped and aligned. The dataset was preprocessed, using the methods described in a previous work~\cite{beckham2017unimodal}, so that the images are 256 pixels in width and height, and pixels values follow a $(0;1)$ normal distribution. The original dataset is split into five cross-validation folds. The training set consists of merging the first four folds which comprise a total of $15554$ images. From this, $10\%$ of the images are held out as part of a validation set. The last fold is used as test set.
	\end{itemize}

	\subsection{The model}
	\label{sect:model}
	Convolutional Neural Networks (CNN) has been used for both datasets. The different architectures of CNN used in these experiments are presented in tables \ref{table:CNNArchitecture} and \ref{table:ResNetArchitecture}. The architecture for DR is the same that was used in \cite{de2018weighted} and the network for Adience is a small Residual Network (ResNet) \cite{he2016deep} that was used in \cite{beckham2017unimodal}. For convolutional layers, ConvWxH@FsS = F filters of size WxH and stride S. For pooling layers, PoolWxHsS = pool size of WxH and stride S.

	\begin{table}[!t]
		\caption{Description of the architecture used in the DR experiments.}
		\label{table:CNNArchitecture}
		\centering
		\begin{tabular}{|l|c|}
			\hline
			\textbf{Layer} & \textbf{Output shape}\\
			\hline
			%Conv3x3@32s1 & 254x254x32\\\hline
			2 x Conv3x3@32s1 & 252x252x32\\\hline
			MaxPool2x2s2 & 126x126x32\\\hline

			%Conv3x3@64s1 & 124x124x64\\\hline
			2 x Conv3x3@64s1 & 122x122x64\\\hline
			MaxPool2x2s2 & 61x61x64\\\hline

			%Conv3x3@128s1 & 59x59x128\\\hline
			2 x Conv3x3@128s1 & 57x57x128\\\hline
			MaxPool2x2s2 & 28x28x128\\\hline

			%Conv3x3@128s1 & 26x26x128\\\hline
			2 x Conv3x3@128s1 & 24x24x128\\\hline
			MaxPool2x2s2 & 12x12x128\\\hline

			Conv4x4@128s1 & 9x9x128\\
			\hline
		\end{tabular}	
	\end{table}

	\begin{table}[!t]
		\caption{Description of the architecture used in the Adience experiments.}
		\label{table:ResNetArchitecture}
		\centering
		\begin{tabular}{|l|c|}
			\hline
			\textbf{Layer} & \textbf{Output shape}\\
			\hline
			Conv7x7@32s2 & 112x112x32\\\hline
			MaxPool3x3s2 & 55x55x32\\\hline
			2 x ResBlock3x3@64s1 & 55x55x32\\\hline
			1 x ResBlock3x3@128s2 & 28x28x64\\\hline
			2 x ResBlock3x3@128s1 & 28x28x64\\\hline
			1 x ResBlock3x3@256s2 & 14x14x128\\\hline
			2 x ResBlock3x3@256s1 & 14x14x128\\\hline
			1 x ResBlock3x3@512s2 & 7x7x256\\\hline
			2 x ResBlock3x3@512s1 & 7x7x256\\\hline
			AveragePool7x7s2 & 1x1x256\\
			\hline
		\end{tabular}	
	\end{table}

>>>>

% !TeX spellcheck = en_GB

@@

% *** PDF, URL AND HYPERLINK PACKAGES ***
%
\usepackage{url}
% url.sty was written by Donald Arseneau. It provides better support for
% handling and breaking URLs. url.sty is already installed on most LaTeX
% systems. The latest version and documentation can be obtained at:
% http://www.ctan.org/pkg/url
% Basically, \url{my_url_here}.

@@

\usepackage{color}

\usepackage{ulem}

\usepackage{array,booktabs}

\usepackage{hyperref}

@@

% correct bad hyphenation here
\hyphenation{op-tical net-works semi-conduc-tor mi-ni-mi-zes des-cri-bed di-ffe-rent sig-ni-fi-can-ce}

@@

	\author{Víctor-Manuel~Vargas, Pedro-Antonio~Gutiérrez and César~Hervás}

@@

	\section{Introduction}
	\label{sect:introduction}
	% The very first letter is a 2 line initial drop letter followed
	% by the rest of the first word in caps.
	% 
	% form to use if the first word consists of a single letter:
	% \IEEEPARstart{A}{demo} file is ....
	% 
	% form to use if you need the single drop letter followed by
	% normal text (unknown if ever used by the IEEE):
	% \IEEEPARstart{A}{}demo file is ....
	% 
	% Some journals put the first two words in caps:
	% \IEEEPARstart{T}{his demo} file is ....
	% 
	% Here we have the typical use of a "T" for an initial drop letter
	% and "HIS" in caps to complete the first word.
	\textcolor{red}{Habría que restructurar la introducción: 1) deep learning y CNN, pero corto 2) clasificación ordinal, modelos de umbral, POM y funciones de enlace (más bien breve) 3) clasificación ordinal profunda 4) objetivo del paper (un estudio experimental sobre funciones de enlace y otros parámetros en deep learning ordinal). learning rate y batch normalization, si se mete, debería ser más breve}

	1) \IEEEPARstart{D}{eep learning}, introduced by Yann Lecun~\cite{lecun2015deep}, combines multiple machine learning techniques and allows computational models that are composed of numerous processing layers to learn representations of data with various levels of abstraction. These methods have dramatically improved the state-of-the-art in many domains, such as image classification~\cite{cirecsan2012multi,he2016deep,krizhevsky2012imagenet}, speech recognition~\cite{hinton2012deep}, control problems~\cite{mnih2015human}, object detection~\cite{jiang2016speed,girshick2014rich}, privacy protection~\cite{yu2017iprivacy}, recovery of human pose~\cite{hong2015multimodal}, semantic segmentation~\cite{long2015fully} and image retrieval~\cite{li2015weakly}. \textcolor{blue}{\sout{Deep learning discovers complex structures in large datasets by using the backpropagation algorithm to change the internal parameters of the model and compute better representations in each layer from the features in the previous layer.} Convolutional Neural Networks (CNN) are one of the types of deep networks that are designed to process data that comes in the form of multiple arrays.} CNNs are appropriate for images, video, speech and audio processing, and they have been used extensively in the last years for automatic classification tasks~\cite{dong2014learning,sun2013deep,ronneberger2015u}. On image classification tasks, each colour channel is represented by a 2D array. In this case, convolutional layers extract the main features from the pixels of the images and, after that, a fully connected layer classify every sample based on its extracted features. At the output of the CNN, a softmax function provides the probabilities of the set of classes predefined in the model for classification tasks. These outputs are compared against the correct values.

	2) Ordinal classification problems are those classification tasks where labels are ordered, and there are different inter-classes importances for each pair of classes. This kind of problem can be treated as a nominal classification problem, but this discards ordinal information. A better approach is to use specific methods that take into account this kind of information to improve the performance of the classification model. \textcolor{blue}{\sout{One way to use the ordinal information is to evaluate the model using an ordinal metric. Multiple metrics exist in the literature of machine learning and statistics \cite{cruz2014metrics,mehdiyev2016evaluating}. Kappa index is a well-known statistic coefficient defined by Cohen \cite{cohen1960coefficient} to measure inter-rater agreement on classifying elements into a set of categories. Later, Weighted Kappa (WK) is a modified version of the Kappa statistic calculated to allow assigning different weights to different levels of aggregation between two variables. An ordinal weighted kappa loss was described in a previous work~\cite{de2018weighted}, which is a differentiable cost function based on the WK metric. This previous work proved that using these functions improves the model performance and reduce the overfitting risk. Different weights can be assigned in an ordinal classification problem. In a linear penalization, the weight is proportional to the distance between the predicted and the real class (in number of categories). In the Quadratic Weighted Kappa (QWK) the penalization is proportional to the square of the distance. In this work, we will use the QWK metric and the QWK loss.} The Proportional Odds Model (POM)~\cite{agresti2010analysis} is an ordinal alternative to the softmax function that was designed for ordinal regression. It belongs to a wider family of models called Cumulative Link Models (CLM). It is inspired in the concept of a latent variable that is projected in an n-dimensional space and a set of thresholds that divides this space into the different ordinal levels. This kind of models uses a link function which can be of different types. The most common link function is the \texttt{logit} one, that is used in POM. However, there are other functions that can be explored. This model will be described in depth in Section \ref{sect:ordinalproblem}.}

	\textcolor{blue}{3) Ordinal models can also be applied to deep learning models. In the case of convolutional networks, the model projection used by the threshold model will be obtained from the last layer of the network. When working with a 1-dimensional space, the last layer will have only one neuron, and its value will be used to classify every sample in the corresponding class according to the thresholds. Some previous works have used the POM in traditional neural networks~\cite{gutierrez2016ordinal}, but it has not been applied to convolutional networks yet. Also, there are some link functions that has not been explored.}

	\textcolor{blue}{4) In this paper, an experimental study regarding link functions for Cumulative Link Models will be made. Also, other parameters that can affect the training process and the model performance, like the learning rate of the optimization algorithm and the batch size, and its interaction will be studied. The nominal version of this model will be used as a baseline for comparison. We will contrast the results obtained with statistical analysis to provide more robust conclusions. An approximated ANOVA III \cite{miller1997beyond} test followed by a posthoc Tukey's test \cite{tukey1949comparing} will be performed because of the limitations of the computational time required to run a higher number of executions.}

	The experiments will be run using two different ordinal datasets: Diabetic Retinopathy~\cite{de2018weighted}, which contains high-resolution fundus images related with diabetes disease, and Adience~\cite{beckham2017unimodal}, which includes human faces images associated with an age range.

	This paper is organized as follows: in Section~\ref{sect:relatedwork}, we take a look at previous works related to this paper. Section~\ref{sect:ordinalproblem} present a formal description of an ordinal problem and the Cumulative Link Models. In Section~\ref{sect:experiments}, we describe the model, the experiments and the datasets used, while, in Section~\ref{sect:results}, we present the results obtained and the statistical analysis. Finally, Section~\ref{sect:conclusions} exposes the conclusions of this work.

%	\textcolor{red}{The backpropagation algorithm is a stochastic gradient descent algorithm that minimises a predefined loss function. It has been used in the last years in many works~\cite{leonard1990improvement,yu1995dynamic,krizhevsky2012imagenet,de2018weighted} for training shallow and deep neural networks. It updates the weights of each layer after backpropagating the loss function gradients through the network. The learning rate hyper-parameter controls the strength of the changes that are applied to the weights. Some works have checked the importance of finding the optimal value for this parameter~\cite{senior2013empirical} and have tried different approaches to improve the training process~\cite{smith2017cyclical}.}
%	
%	\textcolor{red}{Batch normalization is another technique used for this kind of networks. It reduces the internal covariate shift by normalizing layer inputs. It was presented in 2015~\cite{ioffe2015batch} and gives a critical enhancement during the training phase. Batch size is another important hyper-parameter that should be adjusted as it affects the weight updates and also the normalization~\cite{keskar2016large}\cite{radiuk2017impact}.}
%	
%	The softmax function is widely used for the output layer in neural networks for classification tasks. It is a simple and efficient function for multi-class problems, but it is not the best when working with ordinal data. In this paper, the Cumulative Link Model (CLM)~\cite{gutierrez2016ordinal,agresti2010analysis} will be used instead of the softmax function. Different link functions will be explored to compare their performance. Also, the influence of other parameters like the learning rate of the optimization algorithm and the batch size will be studied. We will contrast the results obtained with statistical analysis to provide more robust conclusions. An approximated ANOVA III \cite{miller1997beyond} test followed by a posthoc Tukey's test \cite{tukey1949comparing} will be performed because of the limitations of the computational time required to run a higher number of executions.

	\section{Related works}
	\label{sect:relatedwork}
	There are many works related to the application and development of CNN models. However, few works are focused on ordinal classification problems.

	J. de la Torre et al.~\cite{de2018weighted} proposed the use of a continuous version of the QWK loss function for the optimization algorithm. They compared this cost function against the traditional log-loss function across three different databases, including the Diabetic Retinopathy database as the most complex one. They proved that their function could improve the results as it reduces overfitting and training time. Also, they checked the importance of hyper-parameter tuning. First, they defined QWK metric as follows:
	\begin{equation}
	\textcolor{blue}{\text{QWK} = 1 - \frac{\sum_{i,j} \omega_{i,j} O_{i,j}}{\sum_{i,j} \omega_{i,j} E_{i,j}},}
	\end{equation}
	\textcolor{blue}{where $\omega$ is the penalization matrix, $O$ is the confusion matrix and $E$ is the normalized outer product between the prediction and the true vector.}

	\textcolor{blue}{Then, they provided the QWK loss ($\text{QWK}_l$) definition below:}
	\begin{equation}
	\textcolor{blue}{\text{QWK}_l = 1 - \frac{\sum_{k=1}^N \sum_{c=1}^C \omega_{t_k, c} P_c(X_k)}{\sum_{i=1}^C \frac{N_i}{N} \sum_{j=1}^C ( \omega_{i,j} \sum_{k=1}^N P_j(X_k) ) },}
	\end{equation}
	\textcolor{blue}{where $X_k$ is the input data of the $k$th sample, $N$ is the number of samples, $N_i$ is the number of samples of the $i$th class and $P_c(X_k)$ is the conditional probability that the $k$th sample belongs to class $c$ given that the true class is $t_k$.}

	\textcolor{blue}{Z. Niu et al. \cite{niu2016ordinal} proposed a learning approach to address ordinal regression problems using convolutional neural networks. They divided the problem into a series of binary classification sub-problems and proposed a multiple output CNN optimization algorithm to collectively resolve these classification sub-problems, taking into account the correlation between them.}

	Christopher Beckham and Christopher Pal~\cite{beckham2017unimodal} proposed a straightforward technique to constrain discrete ordinal probability distributions to be unimodal, via the use of the Poisson and binomial probability distributions. They evaluated this approach in the context of deep learning on two large ordinal image datasets, including the Adience dataset used in this paper, obtaining promising results.	\textcolor{blue}{Also, they proposed a simple squared-error reformulation \cite{beckham2016simple} that was sensitive to class ordering.}

	Adience dataset has been used in other works for human age estimation. E. Eidinger \cite{eidinger2014age} presented an approach using support vector machines and neural networks. J.-C. Chen \cite{chen2016cascaded} proposed a coarse-to-fine strategy for deep convolutional networks. G. Levi \cite{levi2015age} presented another convolutional network model for age estimation. \textcolor{red}{Estos son ordinales?} \textcolor{blue}{No son ordinales pero los cito aqui porque luego los uso para comparar con ellos ya que son los que he encontrado que utilicen las base de datos Adience para predecir edad.}

	\textcolor{blue}{H. Li et al. \cite{li2017deep} applied deep learning techniques for solving the ordinal problem of Alzheimer's diagnosis and detecting the different levels of the disease.}

	\textcolor{blue}{Y. Liu et al. \cite{liu2017deep} proposed a new approach of which transforms the ordinal regression problem to binary classification problems and use triplets with instances from different categories to train deep neural networks. In this way, high-level features describing the ordinal relationship are extracted automatically.}

	\textcolor{blue}{A. Rios et al. \cite{rios2017ordinal} presented a CNN model designed to handle ordinal regression tasks on psychiatric notes. They combined an ordinal loss function, a CNN model and conventional feature engineering. Also, they applied a technique called Locally Interpretable Model-agnostic Explanation (LIME) to make the non-linear model more interpretable.}

	\textcolor{blue}{S. Chen et al. \cite{chen2017using} proposed a deep method termed Ranking-CNN. This method combines multiple binary CNNs that are trained with ordinal age labels. The binary outputs are aggregated for the final age prediction and they achieved a tighter error bound for ranking-based age estimation.}

	\textcolor{blue}{H. Fu et al. \cite{fu2018deep} applied deep learning techniques to Monocular Depth Estimation. They introduced a spacing-increasing discretization strategy to treat the problem as an ordinal regression problem. They improved the performance when training the network with an ordinary regression loss. Also, they used a multi-scale network structure that avoids unnecessary spatial pooling.}

	\textcolor{blue}{Y. Liu et al. \cite{liu2018constrained} proposed a constrained optimization formulation for the ordinal regression problem which minimizes de negative loglikelihood for multiple categories constrained by the order relationship between instances.}

	\textcolor{blue}{A. Pal et al. \cite{pal2018severity} defined a loss function for CNN that is based on the Earth Mover's Distance and takes into account the ordinal class relationships.}

	\textcolor{blue}{M. ALALI et al. \cite{alali2018multi} proposed a complex CNN architecture for solving Twitter Sentiment Classification as an ordinal problem. They checked that using average pooling preserves significant features that provide more expresiveness to ordinal scale.}

	\textcolor{blue}{\sout{Currently, there is a great discussion in terms of finding the best activation function, offering good performance and mitigating the effects of the gradient vanishing problem \cite{bengio1994learning,pascanu2013difficulty}. Rectified Linear Unit (ReLU)~\cite{nair2010rectified} is widely used in most deep learning works, but recently, Clevert et al. proposed the Exponential Linear Unit (ELU)~\cite{clevert2015fast}. They proved that ELUs alleviate the vanishing gradient problem via the identity for positive values. In their experiments, ELUs lead not only to faster learning but also to significantly better generalization performance than ReLUs and Leaky ReLU functions (LReLU) on networks with more than five layers. In this way, this paper considers the ELU function for the different experiments.}}

	\textcolor{blue}{\sout{Sergey Ioffe and Christian Szegedy~\cite{ioffe2015batch} described Batch Normalization and its benefits. It reduces the internal covariate shift by normalizing layer inputs. Their method draws its strength from making normalization a part of the model architecture and performing the normalization for each training batch. It allows them to use higher learning rates and be less careful about initialization. It also eliminates the need for using regularization techniques like Dropout.}}

	\section{Cumulative Link Models (CLM)}
	\label{sect:ordinalproblem}
	An ordinal classification problem consists of predicting the label $y$ of an input vector $\mathbf{x}$, where $\mathbf{x} \in \mathcal{X} \subseteq \mathds{R}^K$ and $y \in \mathcal{Y}~=~\{\mathcal{C}_1, \mathcal{C}_2, ..., \mathcal{C}_Q\}$, i.e. $\mathbf{x}$ is in a K-dimensional input space, and $y$ is in a label space of $Q$ different labels. The objective of the ordinal problem is to find a function $r : \mathcal{X} \rightarrow \mathcal{Y}$ to predict the labels or categories of new patterns, given a training set of $N$ points, $D = \{(\mathbf{x}_i, y_i), i = 1, ..., N\}$. Labels have a natural ordering in ordinal problems: $\mathcal{C}_1 \prec \mathcal{C}_2 \prec ... \prec \mathcal{C}_Q$. The order between labels makes it possible to compare two different elements of $\mathcal{Y}$ by using the relation $\prec$. This is not possible under the nominal classification setting. In regression (where $y \in \mathds{R}$), real values in $\mathds{R}$ can be ordered by the standard $<$ operator, but labels in ordinal regression ($y \in \mathcal{Y}$) do not carry metric information, so the category serves as a qualitative indication of the pattern rather than a quantitative one.

	The Proportional Odds Model (POM) arises from a statistical background and is one of the first models designed explicitly for ordinal regression~\cite{mccullagh1980regression}. It dated back to 1980 and is a member of a wider family of models recognised as Cumulative Link Models (CLM)~\cite{agresti2010analysis}. CLMs predict probabilities of groups of contiguous categories, taking the ordinal scale into account. In this way, cumulative probabilities $P(y \prec C_q |\mathbf{x})$ are estimated, which can be directly related to standard probabilities:
	\begin{equation}
		P(y \preceq \mathcal{C}_q | \mathbf{x}) = P(y = \mathcal{C}_1 | \mathbf{x}) + ... + P(y = \mathcal{C}_q | \mathbf{x}),
	\end{equation}
	\begin{equation}
		P(y = \mathcal{C}_q | \mathbf{x}) = P(y \preceq \mathcal{C}_q | \mathbf{x}) - P(y \preceq \mathcal{C}_{q-1} | \mathbf{x}),
	\end{equation}
	with $q = 2, ..., Q-1$, and considering that $P(y = \mathcal{C}_1 | \mathbf{x}) = P(y \preceq \mathcal{C}_1 | \mathbf{x})$ and $P(y \preceq \mathcal{C}_Q | \mathbf{x}) = 1$.

	\textcolor{blue}{The model is inspired by the notion of a latent variable, where $f(x)$ represents a one-dimensional mapping obtained from the output of the last layer, which has only one neuron.} The decision rule $r: \mathcal{X} \rightarrow \mathcal{Y}$ is not fitted directly, but stochastic ordering of space $\mathcal{X}$ is satisfied by the following general model form \cite{herbrich2000large}:
	\begin{equation}
		g^{-1}(P(y \preceq \mathcal{C}_q | \mathbf{x})) = b_q - \textcolor{blue}{f(\mathbf{x})}, \quad q = 1, ..., Q-1,
	\end{equation}
	where $g^{-1} : [0,1] \rightarrow (-\infty, +\infty)$ is a monotonic function often termed as the inverse link function, and $b_q$ is the threshold defined for class $\mathcal{C}_q$. Consider the latent variable $y^* = f(\mathbf{x})^* = \textcolor{blue}{f(\mathbf{x})} + \epsilon$, where $\epsilon$ is the random variable of the error. The most common choice for the probability distribution of $\epsilon$ is the logistic function (which is the default function for POM). Label $\mathcal{C}_q$ is predicted if and only if $f(\mathbf{x}) \in [b_{q-1}, b_q]$, where the function $f$ and $\mathbf{b} = (b_0, b_1, ..., b_{Q-1}, b_Q)$ are to be determined from the data. It is assumed that $b_0 = -\infty$ and $b_Q = +\infty$, so the real line defined by $f(x), x \in \mathcal{X}$, is divided into $Q$ consecutive intervals. Each interval corresponds to a category. The constraints $b_1 \le b_2 \le ... \le b_{Q-1}$ ensures that $P(y \preceq \mathcal{C}_q | \mathbf{x})$ increases with $q$ \cite{mccullagh1980regression}.
	\textcolor{blue}{This order is achieved by defining the first threshold and calculating the rest of thresholds from the first in the following form:}
